Efficient and Scalable Bipartite Matching with Fast Beta Linkage (fabl)

### 2022-11-23 ba_resubmit/response_to_reviewers.tex

	\section*{Reviewer 3}
	\setcounter{responsectr}{0}

	The paper presents a computational variant of the method introduced by Sadinle (2017). Simulation results are presented and the variant is applied on a new, significantly larger data set. A significant part of the paper is a reproduction of the paper of Sadinle. The variant is introduced in section 4. It is simple but consequential. The simulations and applications show how the variant accelerates the computation relative to the original method and reveals the trade-offs.

	The value of the paper resides in that computational complexity continues to be a difficult obstacle in the application of Bayesian statistics. The proposals of the paper enable the application of Sadinle’s approach to an extent not possible before. It is important to present the research for what it is: An extension of the method of Sadinle motivated by computational considerations. Here Sadinle supersedes Fellegi-Sunter in the sense while the starting point of Sadinle’s research is Fellegi-sunter, the starting point of this paper is Sadinle’s work. This should be made clearer, beginning in the abstract and reiterated in the conclusion.

	\reply{%
	Thank you for your positive feedback regarding our paper. We have clarified that our work is an extension of Sadinle (2017) for enhanced speed and scalability. We have tried to make this more clear in both the abstract, introduction, and conclusion. We have also condensed the discussion of Fellegi and Sunter and moved the explanation of the Bayes estimate to the Appendix; we hope that these changes allow the reader to more clearly see our contributions. }

### 2022-11-23 ba_resubmit/kundinger_resubmit.tex
submitted revision

\begin{abstract}
%Recently, researchers have developed Bayesian versions of the Fellegi-Sunter model for record linkage. These have the crucial advantage of quantifying uncertainty from imperfect linkages.  
Within the field of record linkage, Bayesian methods have the crucial advantage of quantifying uncertainty from imperfect linkages. However, current implementations of Bayesian Fellegi-Sunter models are computationally intensive, making them challenging to use on larger-scale record linkage tasks. To address these computational considerations, we propose fast beta linkage (\texttt{fabl}), an extension to the Beta Record Linkage (\texttt{BRL}) method of \cite{sadinle_bayesian_2017}. Specifically, we use independent prior distributions over the matching space, allowing us to use hashing techniques that reduce computational overhead. This also allows us to complete pairwise record comparisons over large datasets through parallel computing and to reduce memory costs through a new technique called storage efficient indexing. Through simulations and two case studies, we show that \texttt{fabl} has markedly increased speed with minimal loss of accuracy when compared to \texttt{BRL}.
%
%\textcolor{blue}{We propose a variation on Bayesian Fellegi-Sunter models that we call fast beta linkage, or \texttt{fabl}, and is motivated by computational considerations.  Our method extends that of \cite{sadinle_bayesian_2017}, who extended the early work of \cite{fellegi_theory_1969}.}

@@

%% ** Keywords **
\begin{keyword}%[class=MSC]
\kwd{data fusion}
\kwd{distributed computing}
\kwd{entity resolution}
\kwd{hashing}
\kwd{record linkage}
\end{keyword}

@@

In many data analysis tasks, analysts seek to identify duplicate records across two data files. This is an increasingly important task in ``data cleaning,'' and is used for inferential and predictive analyses in fields such as statistics, computer science, machine learning, political science, economics, precision medicine, official statistics, and others \citep{christen_2012, gutman2013bayesian, DalzellReiter18, tang2020}. In this article, we consider bipartite record linkage, which merges two data files that contain duplications across, but not within, the respective data files.

Many statistical record linkage methods are extensions of the seminal work of \cite{fellegi_theory_1969} and \cite{newcombe_automatic_1959}. Specifically, \cite{fellegi_theory_1969} created comparison vectors for each pair of records in the data and independently classified each pair as a match or a non-match using a likelihood ratio test. Recent work in the statistical literature has extended this approach for a wide variety of applications \citep{Winkler1990, fair2004generalized, wagner2014person, gill2003english, enamorado2019using, aleshinguendel2021multifile}. Additionally, some methods directly model the linkage variables \citep{steorts_bayesian_2016, marchant_distributed_2019, betancourt2021prior}, but in this paper, we build on the contributions to the comparison vector approach.

@@

In this paper, we propose fast beta linkage (\texttt{fabl}), which extends the \texttt{BRL} model for increased efficiency and scalability. Following the suggestion in \cite{wortman2019}, we relax the one-to-one matching requirement of \texttt{BRL} and use independent priors over the matching space. This allows us to (1) employ hashing techniques that speed up calculations and reduce computational costs, (2) compute the pairwise record comparisons over large datasets via parallel computing, and (3) reduce memory costs through what we call storage efficient indexing. These contributions allow \texttt{fabl} to perform record linkage on much larger data files than previous Bayesian Fellegi-Sunter models at significantly increased speed with minimal loss of accuracy. In particular, computation time under \texttt{BRL} grows quadratically, with the size of each data file, while computation time under \texttt{fabl} grows linearly, only with the size of the smaller data file.

@@

In what follows, Section \ref{sec:review-of_prior-work} reviews the work of \cite{fellegi_theory_1969} and \cite{sadinle_bayesian_2017}. Section \ref{sec:fast-beta-linkage} proposes the \texttt{fabl} model, provides the Gibbs sampler for posterior inference, and shows the loss function used to calculate the Bayes estimate for the bipartite matching. Section \ref{sec:efficiency} introduces the hashing technique and storage efficient indexing used to increase the speed of calculations and the scale of linkage tasks amenable to \texttt{fabl}. Sections \ref{sec:simulations} and \ref{sec:case-studies} demonstrate the speed and accuracy of \texttt{fabl} through simulation studies and case studies of homicides from the El Salvadoran Civil War and the National Long Term Care Study.

@@

Consider two data files $\bm{X}_1$ and $\bm{X}_2$ containing records $\{x_{1i}\}_{i=1}^{n_1}$ and $\{x_{2j}\}_{j=1}^{n_2}$ respectively. Without loss of generality, denote files such that $n_1 \geq n_2$. For ease of readability, we follow the convention established by \cite{sadinle_bayesian_2017} and say ``record $i \in \bm{X}_1$" rather than the more compact $x_{1i}$. Under bipartite matching, the set of matches across data files can be represented in two equivalent ways. First, we may use a matrix $\Delta \in \{0, 1\}^{n_1 \times n_2}$, where
\begin{align}
	\Delta_{ij} =
	\begin{cases}
		1, & \text{if records}\;  i \in \bm{X}_1 \; \text{and}\; j\in \bm{X}_2 \; \text{refer to the same entity}; \\
		0, & \text{otherwise}.\\
	\end{cases}
\end{align}
Though intuitive, this sparse matrix representation can become cumbersome for large linkage tasks. More compactly, bipartite matching also can be viewed as a labeling $\bm{Z} = (Z_1, \ldots, Z_{n_2})$ for the records in $\bm{X}_2$ such that 
\begin{align}
	Z_{j} =
	\begin{cases}
		i, & \text{if records}\;  i \in \bm{X}_1 \; \text{and}\; j\in \bm{X}_2 \; \text{refer to the same entity}; \\
		n_1 + j, & \text{if record}\;  j \in \bm{X}_2 \; \text{does not have a match in}\; \bm{X}_1. \\
	\end{cases}
\end{align}
We can go back and forth between the two using $\Delta_{ij} = I(Z_j = i),$ where $I(\cdot) = 1$ when the expression inside the parentheses is true, and $I(\cdot) = 0$ otherwise.

@@

%Frequently, $\bm{\gamma}_{ij}$ consists of one comparison for each feature shared between the two datasets.

@@

The seminal \cite{fellegi_theory_1969} model employs two independence assumptions: first, that comparison vectors are conditionally independent given their matching status, and second, that the matching status of the record pairs are independent. Using these assumptions, \cite{winkler_state_1999, jaro1989, larsen_2001, enamorado2019using} and others model the comparison data through mixture models of the form
\begin{align}
	\label{eqn:fs_model}
	&\Gamma_{ij} = \gamma_{ij} \mid \Delta_{ij} = 1 \stackrel{iid}{\sim} \mathcal{M}(\bm{m}), \\
	&\Gamma_{ij} = \gamma_{ij} \mid \Delta_{ij} = 0  \stackrel{iid}{\sim} \mathcal{U}(\bm{u}), \notag\\
	& \Delta_{ij}   \stackrel{iid}{\sim} \text{Bernoulli}(\lambda). \notag
\end{align}
Here, $\mathcal{M}$ and $\mathcal{U}$ are the distributions for matching and non-matching record pairs, $\bm{m}$ and $\bm{u}$ are their respective sets of parameters, and $\lambda$ is the marginal probability that a record pair is a match. When using comparison vectors with discrete agreement levels, $\mathcal{M}$ and $\mathcal{U}$ are collections of independent multinomial distributions for each linkage feature. Accordingly, $\bm{m} = (\bm{m}_1, \ldots, \bm{m}_F)$, where $\bm{m}_f = (m_{f1}, \ldots, m_{fL_f})$ and $m_{fl} = P(\gamma_{ij}^f = l|\Delta_{ij} = 1)$ for all fields $f$ and agreement levels $l$. The $\bm{u}$ parameters are defined similarly, with $u_{fl} = P(\gamma_{ij}^f = l|\Delta_{ij} = 0)$.

@@

Within this framework, record pairs are independently classified as matches and non-matches based on the estimated parameters. However, such independent classifications often leads to links that violate one-on-one matching assumptions, requiring post-processing to achieve desirable results. To address this issue, \cite{jaro1989} proposed an optimization technique using the estimated model parameters to produce a bipartite matching. \cite{sadinle_bayesian_2017} later showed that this method could be thought of as a maximum likelihood estimate.

@@

To estimate a one-to-one matching without using a post-processing step, \cite{sadinle_bayesian_2017} incorporated this constraint directly into the model through a Bayesian framework. In addition to standard Dirichlet priors for the $\bm{m}_f$ and $\bm{u}_f$ parameters, he proposed the ``beta distribution for bipartite matching" for the linkage parameter $\bm{Z}$. He assigns a prior distribution for the probability of the indicator that a record in $\bm{X}_2$ has a match in $\bm{X}_1$, so that $I(Z_j \leq n_1) \sim \text{Bernoulli}(\pi)$, where $\pi$ itself is taken to be distributed $\text{Beta}(\alpha_{\pi}, \beta_{\pi})$. It follows that the number of records in $\bm{X}_2$ that have matches, denoted $n_{12}(\bm{Z}) = \sum_{j=1}^{n_2} I(Z_j \leq n_1 + j)$, is distributed according to a $\text{Beta-Binomial}(n_2, \alpha_{\pi}, \beta_{\pi})$. Conditioning on the set of records in $\bm{X}_2$ that have matches, formally denoted $\{I(Z_j \leq n_1)\}_{j=1}^{n_2}$, all $n_1 ! / (n_1 - n_{12}(\bm{Z}))!$ bipartite matchings are taken to be equally likely. Thus, the beta distribution for bipartite matching, is given by
\begin{align}
\label{eqn:sadinle_prior}
P(\bm{Z}|\alpha_{\pi}, \beta_{\pi}) = \frac{(n_1 - n_{12}(\bm{Z}))!}{n_1 !}\frac{\text{B}(n_{12}(\bm{Z}) + \alpha_{\pi}, n_2 - n_{12}(\bm{Z}) + \beta_{\pi})}{\text{B}(\alpha_{\pi}, \beta_{\pi})},
\end{align}
where $\text{B}(\cdot, \cdot)$ represents the Beta function. This prior strictly enforces one-to-one matching, inducing a Gibbs sampler that removes previously matched records from the set of candidate records when sampling each \(Z_j\). This makes the sampler inherently serial, which can be slow when working on linkage tasks with more than a few thousand records.

@@

In the \cite{wortman2019} flat prior, each value $\{1, \ldots, n_1, n_1 +j\}$ is a priori equally likely for $Z_j$. This however amounts to a prior probability of $n_1 / (n_1 + 1)$ that record $j$ has a match in $\bm{X}_1$. In our simulation studies, the flat prior results in poor precision; hence, we prefer (\ref{eqn:fast_beta_prior}). We also note that the flat prior is equivalent to a special case of the fast Beta prior with $\pi$ fixed at the mean of a $\text{Beta}\left(1, 1 / n_1 \right)$ random variable.

@@

\begin{table}[t!]
	\centering
	\begin{tabular}[t!]{ll}
		Symbol & Description \\
		\hline
		$\bm{X}_1, \bm{X}_2$ & datasets\\
		$i \in 1, \ldots, n_1 $ & index over records in $\bm{X}_1$\\
		$j \in 1, \ldots, n_2 $ & index over records in $\bm{X}_2$\\
		$f \in 1, \ldots F$ & index over fields used for comparisons \\
		$l \in 1, \ldots L_f$ & index over agreement levels for feature $f$ \\
		$n_{12}$ & number of entities in common between $\bm{X}_1$ and $\bm{X}_2$\\
		$\bm{\gamma}_{ij}$ & comparison vector for records $i \in \bm{X}_1$ and $j \in \bm{X}_2$ \\
		$Z_j = i$ & records $i \in \bm{X}_1$ and $j \in \bm{X}_2$ match \\
		$Z_j = n_1 + j$ & record $j \in \bm{X}_2$ has no match in $\bm{X}_1$ \\
		$m_{fl}$ & $P(\gamma_{ij}^f = l | Z_j = i)$ \\
		$u_{fl}$ & $P(\gamma_{ij}^f = l | Z_j \neq i)$ \\
		$\pi$ & probability that a record in $\bm{X}_2$ has a match in $\bm{X}_1$ \\
		\hline
	\end{tabular}\caption{Summary of model notation}\label{table_notation_1}
\end{table}

@@

We initialize $\bm{m}$ and $\bm{u}$ from random draws from their prior distributions, and initialize $\bm{Z}$ to reflect no matches across datasets; that is, $\bm{Z} = (n_1 + 1, \ldots, n_1 + n_2)$. To take the $(s+1)$th sample of each $\bm{m}_f$ and $\bm{u}_f$ given $\bm{Z}^s$, we use the full conditionals,
\begin{subequations}
\begin{align}
	\bm{m}_f^{(s+1)}|\Gamma, \bm{Z}^{(s)} &\sim \text{Dirichlet}(\alpha_{f1}(\bm{Z}^{(s)}), \ldots, \alpha_{fL_f}(\bm{Z}^{(s)})), \label{eqn:m_update} \\
	\bm{u}_f^{(s+1)}|\Gamma, \bm{Z}^{(s)} &\sim \text{Dirichlet}(\beta_{f1}(\bm{Z}^{(s)}), \ldots, \beta_{fL_f}(\bm{Z}^{(s)})), \label{eqn:u_update}
\end{align}
\end{subequations}
where $\alpha_{fl}(\bm{Z}^{(s)})= \alpha_{fl} + \sum_{i,j} I_{obs}(\gamma_{ij}^f) I(\gamma_{ij}^f = l) I(z_j^{(s)} = i)$, and $\beta_{fl}(\bm{Z}^{(s)})= \\ \beta_{fl} + \sum_{i,j} I_{obs}(\gamma_{ij}^f) I(\gamma_{ij}^f = l) I(z_j^{(s)} \neq i)$.

@@

We next introduce storage efficient indexing (SEI), which allows us to compute $\mathcal{N}$ and $\mathcal{H}$ for all $n_1 \times n_2$ record pairs, but greatly reduce the memory costs associated with unlikely matches. This allows all-to-all comparisons for substantially larger linkage tasks. All records $i \in \bm{X}_1$ that share agreement pattern $p$ with record $j \in \bm{X}_2$ have the same $w_{ij}$. Therefore, these records have the same probability to be identified as the link for record $j$. Thus, we know that records $i \in r_{j_p}$ such that $N_{j_p}$ is large are very unlikely to be sampled consistently enough to be deemed a match through the Bayes estimate. We know this regardless of the form of the agreement pattern itself, or its associated probabilities. Therefore, rather than store all of these record labels, we store only a small number \(S\) of them in a new nested list $\mathcal{R}^{\text{SEI}}$. Instead of storing \(n_1 \times n_2\) record labels, with SEI we store at most \(n_2 \times P \times S\) labels, regardless of how large \(n_1\) might be. In practice, we recommend $S=10$, as this reduces the number of stored indices for highly unlikely record pairings, but is not likely to eliminate any of the indices for plausible matches. Choosing $S$ too low, like $S=1$ or $S=2$, can concentrate undue mass on few matches and distort linkage results. Posterior calculations still attribute the appropriate weight to all records through the summary statistics in $\mathcal{H}$, and thus we can proceed with posterior inference through the memory reduced $\tilde{\Gamma}^{\text{SEI}} = \{\mathcal{P}, \mathcal{R}^{\text{SEI}}, \mathcal{N}, \mathcal{H}\}$.

%For linkage tasks with large amounts of records, we can partition the two data files \(\bm{X}_1\) and \(\bm{X}_2\) into $C$ and $D$ smaller chunks $\{\bm{X}_{1c} | c = 1, \ldots, C \}$ and $\{\bm{X}_{1d} | d = 1, \ldots, D \}$ for more manageable computations. For each chunked data file $\bm{X}_{1c}$, we conduct all-to-all comparisons with each chunked data file $\bm{X}_{2d}$ to construct the chunked comparison matrix $\Gamma^{cd}$. We then conduct hashing, compress information through SEI, and delete those comparison vectors from memory before continuing with the next chunk of data. These all-to-all comparisons can be computed serially or in parallel. Summary statistics from each pairwise chunk comparison can be easily synthesized to recover sufficient statistics for the full comparison matrix $\Gamma$. Thus, the combination of partitioning, hashing, and SEI allows us to conduct linkage tasks over much larger datasets.

@@

%	We simulate these data for different values of \(n_1\) and \(n_2\), always with $n_2/2$ records in common across datasets. and compare the run-time of \texttt{fabl} against \texttt{BRL}. Since we have 5 binary comparison fields, the number is unique patterns \(P\) is bounded above by \(2^5 = 32\), a bound which is consistently attained in simulations with more records.

@@

	The implementation of \texttt{BRL} that we use is coded in C \citep{sadinle_bayesian_2017}.  In contrast, we use non-optimized code written in \texttt{R} for \texttt{fabl}.  While this complicates comparisons, and indeed disfavors \texttt{fabl}, the computational speed gains for \texttt{fabl} are still evident, especially for larger sample sizes.  Additionally, although \texttt{fabl} is amenable to parallelization, this simulation is run on a single core. Implementing \texttt{fabl} in C++ with paralellization for the hashing step and sampling the matching status of the record pairs should lead to even more computational gains.

@@

\begin{figure}[t]
	\begin{center}
		\includegraphics[width=0.6\textwidth]{../notes/figures/sadinle_sim_plot_partial_DR} 
		\caption{Negative predictive value (NPV), positive predictive value (PPV), and decision rate (DR) on datasets in the simulation in Appendix \ref{partial}. We see poorer performance for \texttt{fabl} only in situations with high overlap.}
		\label{fig:sadinle_simulation_partial}
	\end{center}
\end{figure}

>>>>

\begin{abstract}
%Recently, researchers have developed Bayesian versions of the Fellegi-Sunter model for record linkage. These have the crucial advantage of quantifying uncertainty from imperfect linkages.  
Within the field of record linkage, Bayesian methods have the crucial advantage of quantifying uncertainty from imperfect linkages. However, current implementations of Bayesian Fellegi-Sunter models are computationally intensive, making them challenging to use on larger-scale record linkage tasks. To address these computational considerations, we propose fast beta linkage (\texttt{fabl}), an extension to the Beta Record Linkage (\texttt{BRL}) method of \cite{sadinle_bayesian_2017}. Specifically, we use independent prior distributions over the matching space, allowing us to use hashing techniques that reduce computational overhead. This also allows us to complete pairwise record comparisons over large data files through parallel computing and to reduce memory costs through a new technique called storage efficient indexing. Through simulations and two case studies, we show that \texttt{fabl} has markedly increased speed with minimal loss of accuracy when compared to \texttt{BRL}.
%
%\textcolor{blue}{We propose a variation on Bayesian Fellegi-Sunter models that we call fast beta linkage, or \texttt{fabl}, and is motivated by computational considerations.  Our method extends that of \cite{sadinle_bayesian_2017}, who extended the early work of \cite{fellegi_theory_1969}.}

@@

%% ** Keywords **
\begin{keyword}%[class=MSC]
\kwd{Bayesian methods}
\kwd{distributed computing}
\kwd{entity resolution}
\kwd{hashing}
\kwd{record linkage}
\end{keyword}

@@

In many data analysis tasks, analysts seek to identify duplicate records across two data files. This is an increasingly important task in ``data cleaning'' and is used for inferential and predictive analyses in fields such as statistics, computer science, machine learning, political science, economics, precision medicine, official statistics, and others \citep{christen_2012, gutman2013bayesian, DalzellReiter18, tang2020}. In this article, we consider bipartite record linkage, which merges two data files that contain duplications across, but not within, the respective data files.

Many statistical record linkage methods are extensions of the seminal work of \cite{fellegi_theory_1969} and \cite{newcombe_automatic_1959}. Specifically, \cite{fellegi_theory_1969} created comparison vectors for each pair of records in the data files and independently classified each pair as a match or a non-match using a likelihood ratio test. Recent work in the statistical literature has extended this approach for a wide variety of applications \citep{Winkler1990, fair2004generalized, wagner2014person, gill2003english, enamorado2019using, aleshinguendel2021multifile}. Additionally, some methods directly model the linkage variables \citep{steorts_bayesian_2016, marchant_distributed_2019, betancourt2021prior}, but in this paper, we build on the contributions to the comparison vector approach.

@@

In this paper, we propose fast beta linkage (\texttt{fabl}), which extends the \texttt{BRL} model for increased efficiency and scalability. Following the suggestion in \cite{wortman2019}, we relax the one-to-one matching requirement of \texttt{BRL} and use independent priors over the matching space. This allows us to (1) employ hashing techniques that speed up calculations and reduce computational costs, (2) compute the pairwise record comparisons over large data files via parallel computing, and (3) reduce memory costs through what we call storage efficient indexing. These contributions allow \texttt{fabl} to perform record linkage on much larger data files than previous Bayesian Fellegi-Sunter models at significantly increased speed with minimal loss of accuracy. In particular, computation time under \texttt{BRL} grows quadratically, with the size of each data file, while computation time under \texttt{fabl} grows linearly, only with the size of the smaller data file.

@@

In what follows, Section \ref{sec:review-of_prior-work} reviews the work of \cite{fellegi_theory_1969} and \cite{sadinle_bayesian_2017}. Section \ref{sec:fast-beta-linkage} proposes the \texttt{fabl} model, provides the Gibbs sampler for posterior inference, and shows the loss function used to calculate the Bayes estimate for the bipartite matching. Section \ref{sec:efficiency} introduces the hashing technique and storage efficient indexing used to increase the speed of calculations and the scale of linkage tasks amenable to \texttt{fabl}. Sections \ref{sec:simulations} and \ref{sec:case-studies} demonstrate the speed and accuracy of \texttt{fabl} through simulation studies and case studies of homicides from the El Salvadoran Civil War and the National Long Term Care Study. Finally, Section \ref{discussion} summarizes our contributions and highlights areas for further research.

@@

Consider two data files $\bm{X}_1$ and $\bm{X}_2$ containing records $\{x_{1i}\}_{i=1}^{n_1}$ and $\{x_{2j}\}_{j=1}^{n_2}$ respectively. Without loss of generality, denote files such that $n_1 \geq n_2$. For ease of readability, we follow the convention established by \cite{sadinle_bayesian_2017} and say ``record $i \in \bm{X}_1$" rather than the more compact $x_{1i}$. Under bipartite matching, the set of matches across data files can be represented in two equivalent ways. First, we may use a matrix $\Delta \in \{0, 1\}^{n_1 \times n_2}$, where
\begin{align}
	\Delta_{ij} =
	\begin{cases}
		1, & \text{if records}\;  i \in \bm{X}_1 \; \text{and}\; j\in \bm{X}_2 \; \text{refer to the same entity}; \\
		0, & \text{otherwise}.\\
	\end{cases}
\end{align}
This sparse matrix representation can become cumbersome for large linkage tasks. More compactly, bipartite matching also can be viewed as a labeling $\bm{Z} = (Z_1, \ldots, Z_{n_2})$ for the records in $\bm{X}_2$ such that 
\begin{align}
	Z_{j} =
	\begin{cases}
		i, & \text{if records}\;  i \in \bm{X}_1 \; \text{and}\; j\in \bm{X}_2 \; \text{refer to the same entity}; \\
		n_1 + j, & \text{if record}\;  j \in \bm{X}_2 \; \text{does not have a match in}\; \bm{X}_1. \\
	\end{cases}
\end{align}
We can go back and forth between the two using $\Delta_{ij} = I(Z_j = i),$ where $I(\cdot) = 1$ when the expression inside the parentheses is true, and $I(\cdot) = 0$ otherwise.

@@

%Frequently, $\bm{\gamma}_{ij}$ consists of one comparison for each feature shared between the two data files.

@@

The seminal \cite{fellegi_theory_1969} model employs two independence assumptions: first, that comparison vectors are conditionally independent given their matching status, and second, that the matching status of the record pairs are independent. Using these assumptions, \cite{winkler_state_1999, jaro1989, larsen_2001, enamorado2019using} and others model the comparison data through mixture models of the form
\begin{subequations}
\begin{align}
	&\Gamma_{ij} = \gamma_{ij} \mid \Delta_{ij} = 1 \stackrel{iid}{\sim} \mathcal{M}(\bm{m}), \label{eqn:fs_model} \\
	&\Gamma_{ij} = \gamma_{ij} \mid \Delta_{ij} = 0  \stackrel{iid}{\sim} \mathcal{U}(\bm{u}), \label{eqn:m_dist}\\
	& \Delta_{ij}   \stackrel{iid}{\sim} \text{Bernoulli}(\lambda). \label{eqn:u_dist}
\end{align}
\end{subequations}
Here, $\mathcal{M}$ and $\mathcal{U}$ are the distributions for matching and non-matching record pairs, $\bm{m}$ and $\bm{u}$ are their respective sets of parameters, and $\lambda$ is the marginal probability that a record pair is a match. When using comparison vectors with discrete agreement levels, $\mathcal{M}$ and $\mathcal{U}$ are collections of independent multinomial distributions for each linkage feature. Accordingly, $\bm{m} = (\bm{m}_1, \ldots, \bm{m}_F)$, where $\bm{m}_f = (m_{f1}, \ldots, m_{fL_f})$ and $m_{fl} = P(\gamma_{ij}^f = l|\Delta_{ij} = 1)$ for all fields $f$ and agreement levels $l$. The $\bm{u}$ parameters are defined similarly, with $u_{fl} = P(\gamma_{ij}^f = l|\Delta_{ij} = 0)$.

@@

Within this framework, record pairs are independently classified as matches and non-matches based on the estimated parameters. However, such independent classifications often leads to links that violate one-on-one matching assumptions, requiring post-processing to achieve desirable results. To address this issue, \cite{jaro1989} proposed an optimization technique using the estimated model parameters to produce a bipartite matching. \cite{sadinle_bayesian_2017} later showed that this method is equivalent to a maximum likelihood estimate.

@@

To estimate a one-to-one matching without using a post-processing step, \cite{sadinle_bayesian_2017} incorporates this constraint directly into the model through a Bayesian framework. In addition to standard Dirichlet priors for the $\bm{m}_f$ and $\bm{u}_f$ parameters, he proposes the ``beta distribution for bipartite matching" for the linkage parameter $\bm{Z}$. He assigns a prior distribution for the probability of the indicator that a record in $\bm{X}_2$ has a match in $\bm{X}_1$, so that $I(Z_j \leq n_1) \sim \text{Bernoulli}(\pi)$, where $\pi$ itself is taken to be distributed $\text{Beta}(\alpha_{\pi}, \beta_{\pi})$. It follows that the number of records in $\bm{X}_2$ that have matches, denoted $n_{12}(\bm{Z}) = \sum_{j=1}^{n_2} I(Z_j \leq n_1 + j)$, is distributed according to a $\text{Beta-Binomial}(n_2, \alpha_{\pi}, \beta_{\pi})$. Conditioning on the set of records in $\bm{X}_2$ that have matches, formally denoted $\{I(Z_j \leq n_1)\}_{j=1}^{n_2}$, all $n_1 ! / (n_1 - n_{12}(\bm{Z}))!$ bipartite matchings are taken to be equally likely. Thus, the beta distribution for bipartite matching, is given by
\begin{align}
\label{eqn:sadinle_prior}
P(\bm{Z}|\alpha_{\pi}, \beta_{\pi}) = \frac{(n_1 - n_{12}(\bm{Z}))!}{n_1 !}\frac{\text{B}(n_{12}(\bm{Z}) + \alpha_{\pi}, n_2 - n_{12}(\bm{Z}) + \beta_{\pi})}{\text{B}(\alpha_{\pi}, \beta_{\pi})},
\end{align}
where $\text{B}(\cdot, \cdot)$ represents the Beta function. This prior strictly enforces one-to-one matching, inducing a Gibbs sampler that removes previously matched records from the set of candidate records when sampling each \(Z_j\). This makes the sampler inherently serial, which can be slow when working on linkage tasks with more than a few thousand records.

@@

In the \cite{wortman2019} flat prior, each value $\{1, \ldots, n_1, n_1 +j\}$ is a priori equally likely for $Z_j$. This however amounts to a prior probability of $n_1 / (n_1 + 1)$ that record $j$ has a match in $\bm{X}_1$. In our preliminary studies, the flat prior results in poor precision; hence, we prefer (\ref{eqn:fast_beta_prior}). We also note that the flat prior is equivalent to a special case of the fast Beta prior with $\pi$ fixed at the mean of a $\text{Beta}\left(1, 1 / n_1 \right)$ random variable.

@@

\begin{table}[t!]
	\centering
	\begin{tabular}[t!]{ll}
		Symbol & Description \\
		\hline
		$\bm{X}_1, \bm{X}_2$ & data files\\
		$i \in 1, \ldots, n_1 $ & index over records in $\bm{X}_1$\\
		$j \in 1, \ldots, n_2 $ & index over records in $\bm{X}_2$\\
		$f \in 1, \ldots F$ & index over fields used for comparisons \\
		$l \in 1, \ldots L_f$ & index over agreement levels for feature $f$ \\
		$n_{12}$ & number of entities in common between $\bm{X}_1$ and $\bm{X}_2$\\
		$\bm{\gamma}_{ij}$ & comparison vector for records $i \in \bm{X}_1$ and $j \in \bm{X}_2$ \\
		$Z_j = i$ & records $i \in \bm{X}_1$ and $j \in \bm{X}_2$ match \\
		$Z_j = n_1 + j$ & record $j \in \bm{X}_2$ has no match in $\bm{X}_1$ \\
		$m_{fl}$ & $P(\gamma_{ij}^f = l | Z_j = i)$ \\
		$u_{fl}$ & $P(\gamma_{ij}^f = l | Z_j \neq i)$ \\
		$\pi$ & probability that a record in $\bm{X}_2$ has a match in $\bm{X}_1$ \\
		\hline
	\end{tabular}\caption{Summary of model notation.}\label{table_notation_1}
\end{table}

@@

We initialize $\bm{m}$ and $\bm{u}$ from random draws from their prior distributions, and initialize $\bm{Z}$ to reflect no matches across data files; that is, $\bm{Z} = (n_1 + 1, \ldots, n_1 + n_2)$. To take the $(s+1)$th sample of each $\bm{m}_f$ and $\bm{u}_f$ given $\bm{Z}^s$, we use the full conditionals,
\begin{subequations}
\begin{align}
	\bm{m}_f^{(s+1)}|\Gamma, \bm{Z}^{(s)} &\sim \text{Dirichlet}(\alpha_{f1}(\bm{Z}^{(s)}), \ldots, \alpha_{fL_f}(\bm{Z}^{(s)})), \label{eqn:m_update} \\
	\bm{u}_f^{(s+1)}|\Gamma, \bm{Z}^{(s)} &\sim \text{Dirichlet}(\beta_{f1}(\bm{Z}^{(s)}), \ldots, \beta_{fL_f}(\bm{Z}^{(s)})), \label{eqn:u_update}
\end{align}
\end{subequations}
where $\alpha_{fl}(\bm{Z}^{(s)})= \alpha_{fl} + \sum_{i,j} I_{obs}(\gamma_{ij}^f) I(\gamma_{ij}^f = l) I(z_j^{(s)} = i)$, and $\beta_{fl}(\bm{Z}^{(s)})= \\ \beta_{fl} + \sum_{i,j} I_{obs}(\gamma_{ij}^f) I(\gamma_{ij}^f = l) I(z_j^{(s)} \neq i)$.

@@

We next introduce storage efficient indexing (SEI), which allows us to compute $\mathcal{N}$ and $\mathcal{H}$ for all $n_1 \times n_2$ record pairs but greatly reduce the memory costs associated with unlikely matches. This allows all-to-all comparisons for substantially larger linkage tasks. All records $i \in \bm{X}_1$ that share agreement pattern $p$ with record $j \in \bm{X}_2$ have the same $w_{ij}$. Therefore, these records have the same probability to be identified as the link for record $j$. Thus, we know that records $i \in r_{j_p}$ such that $N_{j_p}$ is large are very unlikely to be sampled consistently enough to be deemed a match through the Bayes estimate. We know this regardless of the form of the agreement pattern itself, or its associated probabilities. Therefore, rather than store all of these record labels, we store only a small number \(S\) of them in a new nested list $\mathcal{R}^{\text{SEI}}$. Instead of storing \(n_1 \times n_2\) record labels, with SEI we store at most \(n_2 \times P \times S\) labels, regardless of how large \(n_1\) might be. In practice, we recommend $S=10$, as this reduces the number of stored indices for highly unlikely record pairings, but is not likely to eliminate any of the indices for plausible matches. Choosing $S$ too low, like $S=1$ or $S=2$, can concentrate undue mass on few matches and distort linkage results. Posterior calculations still attribute the appropriate weight to all records through the summary statistics in $\mathcal{H}$, and thus we can proceed with posterior inference through the memory reduced $\tilde{\Gamma}^{\text{SEI}} = \{\mathcal{P}, \mathcal{R}^{\text{SEI}}, \mathcal{N}, \mathcal{H}\}$.

%For linkage tasks with large amounts of records, we can partition the two data files \(\bm{X}_1\) and \(\bm{X}_2\) into $C$ and $D$ smaller chunks $\{\bm{X}_{1c} | c = 1, \ldots, C \}$ and $\{\bm{X}_{1d} | d = 1, \ldots, D \}$ for more manageable computations. For each chunked data file $\bm{X}_{1c}$, we conduct all-to-all comparisons with each chunked data file $\bm{X}_{2d}$ to construct the chunked comparison matrix $\Gamma^{cd}$. We then conduct hashing, compress information through SEI, and delete those comparison vectors from memory before continuing with the next chunk of data. These all-to-all comparisons can be computed serially or in parallel. Summary statistics from each pairwise chunk comparison can be easily synthesized to recover sufficient statistics for the full comparison matrix $\Gamma$. Thus, the combination of partitioning, hashing, and SEI allows us to conduct linkage tasks over much larger data files.

@@

%	We simulate these data for different values of \(n_1\) and \(n_2\), always with $n_2/2$ records in common across data files. and compare the run-time of \texttt{fabl} against \texttt{BRL}. Since we have 5 binary comparison fields, the number is unique patterns \(P\) is bounded above by \(2^5 = 32\), a bound which is consistently attained in simulations with more records.

@@

	The Gibbs sampler in the implementation of \texttt{BRL} that we use is coded in C \citep{sadinle_bayesian_2017}. In contrast, we use non-optimized code written only in \texttt{R} for \texttt{fabl}.  While this complicates comparisons, and indeed disfavors \texttt{fabl}, the computational speed gains for \texttt{fabl} are still evident, especially for larger sample sizes.  Additionally, although \texttt{fabl} is amenable to parallelization, this simulation is run on a single core. Implementing \texttt{fabl} in C++ with paralellization for the hashing step and sampling the matching status of the record pairs should lead to even more computational gains.

@@

\begin{figure}[t]
	\begin{center}
		\includegraphics[width=0.6\textwidth]{../notes/figures/sadinle_sim_plot_partial_DR} 
		\caption{Negative predictive value (NPV), positive predictive value (PPV), and decision rate (DR) on data files in the simulation in Appendix \ref{partial}. We see poorer performance for \texttt{fabl} only in situations with high overlap.}
		\label{fig:sadinle_simulation_partial}
	\end{center}
\end{figure}
